---
title: "Introduction to snaplocs"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to snaplocs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", fig.path = "man/figures/README-",
  message = FALSE, warning = FALSE, error = FALSE, tidy = TRUE
)
```

`snaplocs` is as package in the functional sector of the SNAPverse. Like most packages, it contains some form of data. However, it is not a data sector package. Data packages are those that contain a large enough amount of data to warrant a standalone package so that functional packages remain small.
This is not one of those packages. Unlike `snappoly` and `snapgrid` it does not contain any maps. `snaplocs` contains point location data in a single data frame. There are 3,914 locations in all, spread across Alaska and western Canada. Compared to SNAPverse data packages, this is quite small.

While `snaplocs` is nevertheless still typically treated as a data package for basic access to the `locs` data set, the package does provide a collection of functions for working with the data. This vignette provides example usage of several functions.

Print a summary of the data. Use the `tibble` package for a concise better print statement.

```{r print}
library(snaplocs)
library(tibble)
locs
```

There are some basic helper functions that grab metadata associated with point locations of interest.

```{r helpers}
x <- "Fairbanks"
get_region(x)
get_country(x)
get_coords(x)
```

These calls also take vectors and `get_coords` with `keep_cols = TRUE` retains all columns of `locs`.

```{r helpers2}
x <- c("Fairbanks", "Calgary")
get_region(x)
get_country(x)
get_coords(x)
get_coords(x, keep_cols = TRUE)
```

Alaska- and western Canada-centric projects at SNAP often use data in a NAD83 Alaska Albers equal area conic projection.
To quickly convert a 2-column matrix or data frame of WGS84 lon/lat coordinates to a 2-column matrix in this projection, do the following:

```{r albers}
wgs2ak(get_coords(x))
head(wgs2ak(locs[3:4]))
```

More functions will be added in the next package revision.
